Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-09 shanghai.tex
Started fixing Frobenius thoughout the paper. Not finished.
Various instances of Frob are screwed up, typographically, in the paper
now. I'll come back to working on them this afternoon and tomorrow.

\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\W}{W}
\newcommand{\Frob}{{\operatorname{Fr}}}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Gr}{Gr}
\DeclareMathOperator{\Irrep}{Irrep}
\DeclareMathOperator{\Pic}{Pic}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Hh}{H}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\Nm}{Nm}

@@

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' here. 
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob_{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob_{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

@@

Throughout this paper, $G$ stands for a commutative group scheme,
smooth and locally of finite type over a finite field $\Fq$; we write
$m : \GxG\to G$ for the multiplication morphism. We fix an
algebraic closure $\bFq$ of $\Fq$ and write $\bG $ for the
commutative group scheme, smooth and locally of finite type over
$\bFq$ obtained by base change; the multiplication morphism for
$\bG$ will be denoted by $\bm$. Let $\Frob_{G}$
(resp. $\Frob_{\GxxG}$) be the Frobenius automorphism of
$\bG$ (resp. $\bGxG$).
\Clifton{I'll insert a reference to SGA where Frobenius is defined for
  a large class of schemes over finite fields.}

@@

%
\labitem{(CS.2)}{CS.2} $F : \Frob_{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob_{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob_{\GxxG}^*\mu}
    && \Frob_{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob_{G}^* \gcs{L} \arrow[swap]{d}{\bm^* F}
    && \Frob_{G}^*\gcs{L}\boxtimes \Frob_{G}^*\gcs{L} \dar{F\boxtimes F} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,F)$ and
  $\cs{L'} = (\gcs{L'},\mu',F')$ are character sheaves on $G$ then
  $\Hom_{\CS(G)}(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob_G^* \gcs{L} \rar{\Frob_G^* \varphi} \arrow[swap]{d}{F} & \Frob_G^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', F\otimes F')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$; {\it cf.} {\tt http://ncatlab.org/nlab/show/Picard+group}.
\Clifton{Replace ncatlab with other references.}

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob_G$ and
therefore an automorphism $F_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically (this
follows from the Condition~\ref{CS.1}). The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $t^\Frob_\cs{L} : G(\Fq)\to \EEx$,
defined by $g \mapsto F_{\bg}$, is a group homomorphism;
we refer to $t^\Frob_\cs{L}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
\Clifton{There is no actual trace involved, since $\gcs{L}$ is a rank
  $1$ local system. More generally, trace is needed.}

It is clear that $t^\Frob_{\cs{L}\otimes \cs{L'}} = t^\Frob_\cs{L} \cdot t^\Frob_\cs{L'}$
(pointwise multiplication of functions) and that $t^\Frob_\cs{L} = t^\Frob_\cs{L'}$
if $\cs{L} \iso \cs{L'}$ in $\CS(G)$. In this way we obtain a group homomorphism
\[
\cdef{t_{G}} : \CSiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto t^\Frob_\cs{L},
\]
from the Picard group of character sheaves on $G$ to the character group of $G(\Fq)$.

@@

  To see that $(\bfxf)^* \gcs{L}$ satisfies
  condition \ref{CS.1} simply apply the functor $(\bfxf)^*$
  to \ref{CS.1} for $\cs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gcs{L}$
  (a few times). Likewise, to see that $\bf^*\gcs{L}$ satisfies condition
  \ref{CS.2}, apply the same functor to \ref{CS.2} for $\gcs{L}$.
  Then use the fact that $f$ is defined over $\Fq$, so
  $(\bfxf)^*\Frob_{\GxxG}^* \iso \Frob_{\GxxG} (\bfxf)^*$,
  and the fact that $f$ is a morphism of group schemes, so
  $(\bfxf)^* \bm^*\iso \bm^* \bf^*$
  (isomorphisms of functors on constructible sheaves).

@@

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, F)$ is a character sheaf on $G$ we may define
a character sheaf $\nu(\cs{L}) = (\gcs{L}, \mu, F_{k'})$ on $G_{k'}$ by setting
\[
F_{k'} = F \circ \Frob_G^*(F) \circ \cdots \circ (\Frob_G^{n-1})^*(F).
\]
The commutativity of the diagram (CS.2) for $F_{k'}$ follows from the fact that $\Frob_{G_{k'}} = \Frob_G^n$.
Note that we may also think about the construction of $F_{k'}$ as taking the action of $\Weil{k}$ defined in Remark
\ref{rem:CS-action} and restricting it to $\Weil{k'}$.

@@

\begin{proof}
 Let $b_G : {\bar G} \to G$ be the pull-back of $\Spec{\bFq} \to \Spec{\Fq}$ along $G\to \Spec{\Fq}$.
 Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf on $G$. 
 Then $\cs{L}_0$ is, in particular, an $\ell$-adic constructible sheaf on $G$. 
 Then $b_G^* \cs{L}_0$ comes equipped with an isomorphism 
 $F : \Frob_G^* b_G^*\cs{L}_0 \to b_G^* \cs{L}_0$; 
 moreover, the functor $\cs{L}_0 \mapsto (b_G^* \cs{L}_0,F)$  
 from $\ell$-adic constructible sheaves on $G$ to $\ell$-adic constructible sheaves on $G$
 is full and faithful; see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}. 
 This functor preserves local constancy, so takes local systems to local systems. 
 Set $\mu = b_{G\times G}^*\mu_0$; clearly, this satisfies Condition~\ref{CS.2} 
 with $b_G^*\cs{L}_0$ playing the role of $\gcs{L}$.
 Moreover, $F$ is compatible with $\mu$ in the sense of Condition~\ref{CS.3}.
 It follows that functor $(\cs{L}_0,\mu_0) \mapsto (b_G^*\cs{L}_0,b_{G\times G}^* \mu_0, F)$, denoted by  $B_G^*: \CSb(G) \hookrightarrow \CS(G)$ below, is full and faithful.
\end{proof}

@@

%\Clifton{I'll address this point in the next section: "Because $\Gal(\bFq/\Fq)$ acts trivially on $\Aut({\bar f})$, the local system ${\bar L}_G(f,\psi)\ceq b_G^* L(f,\psi)$ comes equipped with a canonical isomorphism $F : \Frob_G^* {\bar L}_G(f,\psi) \to {\bar L}_G(f,\psi)$ of constructible sheaves on ${\bar G}$."}

@@

  Let $g$ be any $\Fq$-rational point on $G$ 
  and let $\bg$ be the geometric point on $G$ lying above $g$ 
  (we fixed $\Fq \hookrightarrow \bFq$ long ago). 
  Since $G$ is connected, the geometric point $\bg$ determines
  an equivalence between the category of $\ell$-adic Weil local systems on $G$ and
  $\ell$-adic representations of $\W(G,\bg)$ \cite{Deligne:Weil2}*{1.1.12}. 
  Now let $(\gcs{L},\mu,F)$ be a character sheaf on $G$ 
  and let $\rho : \W(G, \bg) \to \EEx$ be the character determined by $(\gcs{L},F)$. 
  The $\Fq$-rational point $g$ under the geometric point $\bg$ determines a splitting
  $\Weil{\Fq}\to \W(G,\bg)$ of $\W(G,\bg)\to \Weil{\Fq}$. 
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character $\rho_g : \Weil{\Fq} \to \EEx$.
  %\[
  %\begin{tikzcd}
  %1 \rar & \pi_1(\bG, \bg) \rar &  \arrow{d}{\rho} \W(G,\bg) \rar
  %& \ar[swap,bend right, dotted]{l}{g} \arrow[dotted]{dl}{\rho_g} \Weil{\Fq} \rar & 1\\
  %&&\EEx&&
  %\end{tikzcd}
  %\]
  This character is given by the trace of Frobenius of $\cs{L}$ at $g$, as defined in Section~\ref{sec:Frob}:
  \[
  \rho_g(\Frob) =  t^\Frob_\cs{L}(g).
  \]

  On the other hand, we have already seen that $t^\Frob_\cs{L} : G(\Fq) \to \EEx$
  is a group homomorphism. Since $G$ is an algebraic group over $\Fq$, it is a
  variety over $\Fq$, so finitely generated over $\Fq$, so $G(\Fq)$ is finite;
  it follows that $t^\Frob_\cs{L}(g) = \rho_g(\Frob)$ is a root of unity,
  for every $g\in G(\Fq)$.  Since $\Weil{\Fq}$ is generated by
  $\Frob$ and $\rho_g : \Weil{\Fq} \to \EEx$ is
  a character, it follows that the image of $\rho_g$ is a finite group.
  Thus, $\rho_g$ extends to an $\ell$-adic character of $\Gal(\bFq/\Fq)$, 
  denoted by the same symbol, below.
  \[
  \begin{tikzcd}
  1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \W(G,\bg) \rar \dar & \Weil{\Fq} \arrow[bend left]{rr}{\rho_g} \rar \dar & 1 & \EEx\\
  1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \arrow[dotted]{rru} \rar & 1 &
  \end{tikzcd}
  \]

@@

  Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf on $G$. 
  Then $\cs{L}\ceq B_G^*(\cs{L}_0,\mu_0)$ is a character sheaf on $G$.
  Recall the definition of the character $t_\cs{L}^\Frob : G(\Fq) \to \EEx$ from Section~\ref{sec:Frob}. 
  Let $f : G\to G$ be the Lang isogeny.
  Recall that $\ker f = G(\Fq) = G^\Frob$; 
  in particular, the action of $\Gal(\bFq/\Fq)$ on $\ker f$ is trivial.
  Let $V = \EE$ and let $\psi : \ker f \to \Aut(V)$ be the representation given by 
  $\psi(g)(v) = t_\cs{L}^\Frob(g) v$ for $v\in V$. 
  Then $(f,\psi)$ is a finite character sheaf.

  To show that $L_G(f,\psi) \iso (\cs{L}_0,\mu_0)$ we simply apply the functions--sheaves dictionary
  as follows ({\it cf} \cite{Laumon}*{1.1.3} or, ultimately, \cite{SGA4.5}*{Sommes trigonométriques}).
  First, recall that $L_G(f,\psi) = (L(f,\psi), \mu(f,\psi))$. 
  By  \cite{Laumon}*{1.1.3.3}, $t_{L(f,\psi)}^{\Frob^n} = t_{\cs{L}}^{\Frob}\circ N_{\FF_{q^n}/\FF_q}$, for each positive integer $n$.
  On the other hand, $t_{\cs{L}}^{\Frob} \circ N_{\FF_{q^n}/\FF_q} = t_{\cs{L}}^{\Frob^n}$, for each positive integer $n$, 
  by Proposition~\ref{prop:basechange}.
  %Using notation from \cite{Laumon}*{1.1}, this means $t_{\bullet L(f,\psi)} = t_{\bullet \cs{L}_0}$.
  Using \cite{Laumon}*{Th\'eor\`eme~1.1.2}, it follows that  $L(f,\psi) \iso \cs{L}_0$ in $\Loc(G)$.
  It is now clear that $(L(f,\psi),\mu(f,\psi)) \iso (\cs{L}_0,\mu_0)$ in $\CSb(G)$.
  This completes the proof that $L_G : \CSf(G) \to \CS(G)$ is essentially surjective.
  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it follows that $L_G$ is an equivalence.

@@

\begin{proposition}\label{prop:etale}
  If $G$ is a commutative \'etale group scheme over $\Fq$ and
  $G(\bFq)$ is finitely generated then there is a canonical
  isomorphism
  \[
  \CSiso{G} \longrightarrow \Hh^1(\Weil{\Fq}, \Hom_\text{grp}(G(\bFq),\EEx)),
  \qquad \text{given by}\qquad
  [\cs{L}] \mapsto [t_\cs{L}],
  \]
  where $t_\cs{L}: \Weil{\Fq}\to \Hom_\text{grp}(G(\bFq),\EEx)$ is a cocycle such that
  \[
  \forall g\in G(\Fq), \qquad t_\cs{L}(\Frob)(g) = t^\Frob_\cs{L}(g).
  \]
\end{proposition}

>>>>

\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\W}{W}
\newcommand{\Frob}[1]{{\operatorname{F}_{#1}}}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Gr}{Gr}
\DeclareMathOperator{\Irrep}{Irrep}
\DeclareMathOperator{\Pic}{Pic}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Hh}{H}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\Nm}{Nm}
\DeclareMathOperator{\trace}{Tr}

@@

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' here. 
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

@@

Throughout this paper, $G$ stands for a commutative group scheme,
smooth and locally of finite type over a finite field $\Fq$; we write
$m : \GxG\to G$ for the multiplication morphism. We fix an
algebraic closure $\bFq$ of $\Fq$ and write $\bG $ for the
commutative group scheme, smooth and locally of finite type over
$\bFq$ obtained by base change; the multiplication morphism for
$\bG$ will be denoted by $\bm$. 
Let $\Frob{G}$ be the Frobenius automorphism of $\bG$.
\Clifton{I was a bit glib about Frobenius, previously.
Now I've made things more precise here; I'm in the process of tracking this improvement through the paper.}
Recall that $\Frob{G} = \id_{G}\times \Frob{\Spec{\bFq}}$ where
$\Frob{\Spec{\bFq}}$ is the automorphism of $\Spec{\bFq}$ 
opposite to the geometric Frobenius element $\Frob{\Fq}$ in $\Weil{\Fq} \subset \Gal(\bFq/\Fq)$.

@@

%
\labitem{(CS.2)}{CS.2} $F : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* F}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes F} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,F)$ and
  $\cs{L'} = (\gcs{L'},\mu',F')$ are character sheaves on $G$ then
  $\Hom_{\CS(G)}(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \varphi} \arrow[swap]{d}{F} & \Frob{G}^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', F\otimes F')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$.\footnote{See {\tt http://ncatlab.org/nlab/show/Picard+group}.}

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $F_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically (this
follows from the Condition~\ref{CS.1}). The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $t^\Frob{\Fq}\cs{L} : G(\Fq)\to \EEx$,
defined by $g \mapsto F_{\bg}$, is a group homomorphism;
we refer to $t^\Frob{\Fq}\cs{L}$ as the \cdef{trace of Frobenius of $\cs{L}$}, 
even though no trace is involved in the definition, because the notion extends 
to any constructible $\ell$-adic Weil sheaf $(\gcs{L},F)$ on $G$, in which context,
$t^\Frob{(\gcs{L},F)} : G(\Fq)\to \EE$ is defined by $t^\Frob{(\gcs{L},F)}(g) \ceq \trace F_{\bg}$.

It is clear that $t^\Frob{\cs{L}\otimes \cs{L'}} = t^\Frob{\Fq}\cs{L} \cdot t^\Frob{\Fq}\cs{L'}$
(pointwise multiplication of functions) and that $t^\Frob{\Fq}\cs{L} = t^\Frob{\Fq}\cs{L'}$
if $\cs{L} \iso \cs{L'}$ in $\CS(G)$. In this way we obtain a group homomorphism
\[
\cdef{t_{G}} : \CSiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto t^\Frob{\Fq}\cs{L},
\]
from the Picard group of character sheaves on $G$ to the character group of $G(\Fq)$.

@@

  To see that $(\bfxf)^* \gcs{L}$ satisfies
  condition \ref{CS.1} simply apply the functor $(\bfxf)^*$
  to \ref{CS.1} for $\cs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gcs{L}$
  (a few times). Likewise, to see that $\bf^*\gcs{L}$ satisfies condition
  \ref{CS.2}, apply the same functor to \ref{CS.2} for $\gcs{L}$.
  Then use the fact that $f$ is defined over $\Fq$, so
  $(\bfxf)^*\Frob{\GxxG}^* \iso \Frob{\GxxG} (\bfxf)^*$,
  and the fact that $f$ is a morphism of group schemes, so
  $(\bfxf)^* \bm^*\iso \bm^* \bf^*$
  (isomorphisms of functors on constructible sheaves).

@@

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, F)$ is a character sheaf on $G$ we may define
a character sheaf $\nu(\cs{L}) = (\gcs{L}, \mu, F_{k'})$ on $G_{k'}$ by setting
\[
F_{k'} = F \circ \Frob{G}^*(F) \circ \cdots \circ (\Frob{G}^{n-1})^*(F).
\]
The commutativity of the diagram (CS.2) for $F_{k'}$ follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $F_{k'}$ as taking the action of $\Weil{k}$ defined in Remark
\ref{rem:CS-action} and restricting it to $\Weil{k'}$.

@@

\begin{proof}
 Let $b_G : {\bar G} \to G$ be the pull-back of $\Spec{\bFq} \to \Spec{\Fq}$ along $G\to \Spec{\Fq}$.
 Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf on $G$. 
 Then $\cs{L}_0$ is, in particular, an $\ell$-adic constructible sheaf on $G$. 
 Then $b_G^* \cs{L}_0$ comes equipped with an isomorphism 
 $F : \Frob{G}^* b_G^*\cs{L}_0 \to b_G^* \cs{L}_0$; 
 moreover, the functor $\cs{L}_0 \mapsto (b_G^* \cs{L}_0,F)$  
 from $\ell$-adic constructible sheaves on $G$ to $\ell$-adic constructible sheaves on $G$
 is full and faithful; see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}. 
 This functor preserves local constancy, so takes local systems to local systems. 
 Set $\mu = b_{G\times G}^*\mu_0$; clearly, this satisfies Condition~\ref{CS.2} 
 with $b_G^*\cs{L}_0$ playing the role of $\gcs{L}$.
 Moreover, $F$ is compatible with $\mu$ in the sense of Condition~\ref{CS.3}.
 It follows that functor $(\cs{L}_0,\mu_0) \mapsto (b_G^*\cs{L}_0,b_{G\times G}^* \mu_0, F)$, denoted by  $B_G^*: \CSb(G) \hookrightarrow \CS(G)$ below, is full and faithful.
\end{proof}

@@

%\Clifton{I'll address this point in the next section: "Because $\Gal(\bFq/\Fq)$ acts trivially on $\Aut({\bar f})$, the local system ${\bar L}_G(f,\psi)\ceq b_G^* L(f,\psi)$ comes equipped with a canonical isomorphism $F : \Frob{G}^* {\bar L}_G(f,\psi) \to {\bar L}_G(f,\psi)$ of constructible sheaves on ${\bar G}$."}

@@

  Let $g$ be any $\Fq$-rational point on $G$ 
  and let $\bg$ be the geometric point on $G$ lying above $g$ 
  (we fixed $\Fq \hookrightarrow \bFq$ long ago). 
  Since $G$ is connected, the geometric point $\bg$ determines
  an equivalence between the category of $\ell$-adic Weil local systems on $G$ and
  $\ell$-adic representations of $\W(G,\bg)$ \cite{Deligne:Weil2}*{1.1.12}. 
  Now let $(\gcs{L},\mu,F)$ be a character sheaf on $G$ 
  and let $\rho : \W(G, \bg) \to \EEx$ be the character determined by $(\gcs{L},F)$. 
  The $\Fq$-rational point $g$ under the geometric point $\bg$ determines a splitting
  $\Weil{\Fq}\to \W(G,\bg)$ of $\W(G,\bg)\to \Weil{\Fq}$. 
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character $\rho_g : \Weil{\Fq} \to \EEx$.
  %\[
  %\begin{tikzcd}
  %1 \rar & \pi_1(\bG, \bg) \rar &  \arrow{d}{\rho} \W(G,\bg) \rar
  %& \ar[swap,bend right, dotted]{l}{g} \arrow[dotted]{dl}{\rho_g} \Weil{\Fq} \rar & 1\\
  %&&\EEx&&
  %\end{tikzcd}
  %\]
  This character is given by the trace of Frobenius of $\cs{L}$ at $g$, as defined in Section~\ref{sec:Frob}:
  \[
  \rho_g(\Frob) =  t^\Frob{\Fq}\cs{L}(g).
  \]

  On the other hand, we have already seen that $t^\Frob{\Fq}\cs{L} : G(\Fq) \to \EEx$
  is a group homomorphism. Since $G$ is an algebraic group over $\Fq$, it is a
  variety over $\Fq$, so finitely generated over $\Fq$, so $G(\Fq)$ is finite;
  it follows that $t^\Frob{\Fq}\cs{L}(g) = \rho_g(\Frob)$ is a root of unity,
  for every $g\in G(\Fq)$.  Since $\Weil{\Fq}$ is generated by
  $\Frob{\Fq}$ and $\rho_g : \Weil{\Fq} \to \EEx$ is
  a character, it follows that the image of $\rho_g$ is a finite group.
  Thus, $\rho_g$ extends to an $\ell$-adic character of $\Gal(\bFq/\Fq)$, 
  denoted by the same symbol, below.
  \[
  \begin{tikzcd}
  1 \rar & \ar[equal]{d} \pi_1(\bG, \bg) \rar & \W(G,\bg) \rar \dar & \Weil{\Fq} \arrow[bend left]{rr}{\rho_g} \rar \dar & 1 & \EEx\\
  1 \rar &  \pi_1(\bG, \bg) \rar & \pi_1(G,\bg) \rar & \Gal(\bFq/\Fq) \arrow[dotted]{rru} \rar & 1 &
  \end{tikzcd}
  \]

@@

  Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf on $G$. 
  Then $\cs{L}\ceq B_G^*(\cs{L}_0,\mu_0)$ is a character sheaf on $G$.
  Recall the definition of the character $t_\cs{L}^\Frob : G(\Fq) \to \EEx$ from Section~\ref{sec:Frob}. 
  Let $f : G\to G$ be the Lang isogeny.
  Recall that $\ker f = G(\Fq) = G^\Frob{\Fq}$; 
  in particular, the action of $\Gal(\bFq/\Fq)$ on $\ker f$ is trivial.
  Let $V = \EE$ and let $\psi : \ker f \to \Aut(V)$ be the representation given by 
  $\psi(g)(v) = t_\cs{L}^\Frob{\Fq}(g) v$ for $v\in V$. 
  Then $(f,\psi)$ is a finite character sheaf.

  To show that $L_G(f,\psi) \iso (\cs{L}_0,\mu_0)$ we simply apply the functions--sheaves dictionary
  as follows ({\it cf} \cite{Laumon}*{1.1.3} or, ultimately, \cite{SGA4.5}*{Sommes trigonométriques}).
  First, recall that $L_G(f,\psi) = (L(f,\psi), \mu(f,\psi))$. 
  By  \cite{Laumon}*{1.1.3.3}, $t_{L(f,\psi)}^\Frob{k'} = t_{\cs{L}}^\Frob{k}\circ N_{k'/k}$, for each finite extension $k'/k$.
  On the other hand, $t_{\cs{L}}^\Frob{k} \circ N_{k'/k} = t_{\cs{L}}^\Frob{k'}$, for each for each finite extension $k'/k$, 
  by Proposition~\ref{prop:basechange}.
  %Using notation from \cite{Laumon}*{1.1}, this means $t_{\bullet L(f,\psi)} = t_{\bullet \cs{L}_0}$.
  Using \cite{Laumon}*{Th\'eor\`eme~1.1.2}, it follows that  $L(f,\psi) \iso \cs{L}_0$ in $\Loc(G)$.
  It is now clear that $(L(f,\psi),\mu(f,\psi)) \iso (\cs{L}_0,\mu_0)$ in $\CSb(G)$.
  This completes the proof that $L_G : \CSf(G) \to \CS(G)$ is essentially surjective.
  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it follows that $L_G$ is an equivalence.

@@

\begin{proposition}\label{prop:etale}
  If $G$ is a commutative \'etale group scheme over $\Fq$ and
  $G(\bFq)$ is finitely generated then there is a canonical
  isomorphism
  \[
  \CSiso{G} \longrightarrow \Hh^1(\Weil{\Fq}, \Hom_\text{grp}(G(\bFq),\EEx)),
  \qquad \text{given by}\qquad
  [\cs{L}] \mapsto [t_\cs{L}],
  \]
  where $t_\cs{L}: \Weil{\Fq}\to \Hom_\text{grp}(G(\bFq),\EEx)$ is a cocycle such that
  \[
  \forall g\in G(\Fq), \qquad t_\cs{L}(\Frob)(g) = t^\Frob{\Fq}\cs{L}(g).
  \]
\end{proposition}
